```{r}
library(tidyverse)
library(MASS)
library(ggplot2)
library(faraway)

data <- read.csv("kc_house_data.csv")
data
```
Jett's Data Cleaning:
```{r}
non_renovated_data <- data[data$yr_renovated == 0, ]
renovated_data <- data[data$yr_renovated != 0, ]

# Katherine's Cleaning
data = data %>% 
  filter(bedrooms != 33)

data1 = data

data1$date = substring(data$date, 1,8)
data1$date = as.Date(strptime(data1$date, "%Y%m%d"))
data1$month = month(ymd(data1$date), label = TRUE, abbr = FALSE)
data1$year = year(data1$date)

Data3 = data1 %>%
  group_by(month)%>%
  summarize(avg_price = mean(price))

data = data %>%
  mutate(good_quality = case_when(
    condition > 3 & grade > 7  ~ 1 ,
    condition <= 3 | grade <= 7 ~ 0))

data$good_quality = as.factor(data$good_quality)

```

```{r}
invalid_renovation_data <- data[data$yr_renovated != 0 & data$yr_renovated < data$yr_built, ]
```


# Code to split our data into test and training
```{r}
set.seed(6021)
sample.data<-sample.int(nrow(data), floor(.50*nrow(data)), replace = F)

train<-data[sample.data, ]
test<-data[-sample.data, ]
```

```{r}

ggplot(data, aes(x = sqft_living, y = price)) +
  geom_point(alpha = 0.4) +
  labs(
    title = "Scatter Plot: Price vs. Living Area",
    x = "Living Area (sqft)",
    y = "Price"
  )
```
```{r}
ggplot(data, aes(x = as.factor(condition), y = price)) +
  geom_boxplot() +
  labs(
    title = "Box Plot: Price by Condition",
    x = "Condition",
    y = "Price"
  )
```

```{r}
ggplot(data, aes(x = as.factor(condition), y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Box Plot: Price by Condition",
    x = "Condition",
    y = "Price (Log)"
  )
```

Zoha Data Visualizations:
```{r}
ggplot(data, aes(x=sqft_living, y=price, color=sqft_living)) +
                 facet_wrap(~waterfront) +
                 geom_point() + 
                 labs(title = "Price of House based on sqft living and waterfront view",
                        x="Square foot living space",
                        y="Price")+ 
                theme(plot.title = element_text(hjust = 0.5))
```
**(2) Price of median sqft living **
```{r}
Data<-data%>%
  mutate(house_size=case_when(sqft_living <= 2000 ~ "small",
  sqft_living > 2000 & sqft_living <=5000 ~ "medium",
  sqft_living > 5000 ~ "large"))

medianPrice<-Data %>% 
  group_by(house_size) %>% 
  summarize(median_price=median(price)) 

ggplot(medianPrice, aes(x=house_size, y=median_price))+
  geom_bar(fill="#90e7fc", stat='identity')+
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 90))+
  labs(x="Square foot living", y="Median Price of House",
  title="Price of House by Sq Ft")
```

**(3) Price and grade **

```{r}
ggplot(Data, aes(x=grade, y=price, color=grade)) +
                 geom_point() + 
                 labs(title = "House Price based on Grade",
                        x="Grade",
                        y="Price") + 
                theme(plot.title = element_text(hjust = 0.5))
```
**(3) Price and grade boxcox plot**

```{r}
ggplot(Data, aes(x= as.factor(grade), y=price))+ 
  geom_boxplot()+
  theme(plot.title = element_text(hjust=0.5))+
  labs(x="Grade", y="Price", title="Price of House across Grade")
```

Katherine's Data Visualizations
```{r}
####### Scatterplot of Price over Bedrooms ########### 
ggplot(data, aes(y=price,x=bedrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bedroom Count",x="Bedroom Count", y="Price ($ USD)") + geom_smooth(method=lm, se=FALSE)


####### Scatterplot of Price over Bathrooms ########### 
ggplot(data, aes(y=price,x=bathrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bathroom Count", x="Bathroom Count", y="Price ($ USD)")+ geom_smooth(method=lm, se=FALSE)

ggplot(data,aes(x=bedrooms,y=price)) + 
        geom_bar(stat = "identity", na.rm=TRUE)


########### Scatterplot of Price Over Time ########### 
ggplot(data, aes(y=price,x=date))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price Over Time",x="Time", y="Price ($ USD)")

########### Histogram Total Price by Month ########### 
ggplot(data,aes(x=month)) + 
        geom_bar(na.rm=TRUE)+
         labs(title = "Histogram: Total Houses Sold Over Time",x="Time", y="Count")


########### Average Price by Month ########### 
ggplot(Data3, aes(y=avg_price,x=month))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Average Price Over Time",x="Month", y="Price ($ USD)")

```

Jett's LM, forward, backward and stepwise elimination:
```{r}
#I removed time due to the difficulties in working with it in a linear regression model, and I took out sqft living as it is just the sum of sqft above and sqft basement and is therefore redundant
result <- lm(price ~ . - date - sqft_living - id, data = data)
summary(result)
```
```{r}
regnull <- lm(price~1, data=data)
regfull <- result
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```
```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")
```
```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```
Jett regression assumptions check:
```{r}
par(mfrow = c(2, 2))
plot(result)
```
```{r}
library(MASS)
MASS::boxcox(result)
```

```{r}
#data$y.star<-log(data$price)
#result.ystar<-lm(y.star~.- date - sqft_living - id, data=data)
#par(mfrow = c(2, 2))
#plot(result.ystar)
```
Jett, Checking for multicollinearity:
```{r}
round(faraway::vif(result),3)
```
No variables have a VIF of above 5, meaning this test does not indicate multicolinearity. sqftabove however has a 4.923 which is close enough to look into!

Jett, Trying out potential other model:
```{r}
reduced <- lm(price ~ . - id - sqft_basement - date, data = data)
reduced
```
```{r}
par(mfrow = c(2, 2))
plot(reduced)
```
```{r}
library(faraway)
round(faraway::vif(reduced),3)
```

Jett, ANOVA test comparison
```{r}
anova(reduced, result)
```
This supports the full model over the reduced model







Katherine's Logistic Model
```{r}
#Correlations of quantitative predictors
round(cor(train[,3:21], use= "complete.obs"),3)

train$good_quality = as.factor(train$good_quality)

dp1<-ggplot2::ggplot(train,aes(x=bedrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bedrooms by House Quality")

dp2<-ggplot2::ggplot(train,aes(x=bathrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bathrooms by House Quality")

dp3<-ggplot2::ggplot(train,aes(x=sqft_living, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bedrooms by House Quality")

dp4<-ggplot2::ggplot(train,aes(x=floors, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Floors by House Quality")

gridExtra::grid.arrange(dp1, dp2, dp3, dp4, ncol = 2, nrow = 2)
```

```{r}

dp5<-ggplot2::ggplot(train,aes(x=sqft_lot, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft of Lot by House Quality")

dp6<-ggplot2::ggplot(train,aes(x=waterfront, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Waterfron by House Quality")

dp7<-ggplot2::ggplot(train,aes(x=view, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of View by House Quality")

dp8<-ggplot2::ggplot(train,aes(x=sqft_above, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Above by House Quality")

gridExtra::grid.arrange(dp5, dp6, dp7, dp8, ncol = 2, nrow = 2)

```

```{r}
dp9<-ggplot2::ggplot(train,aes(x=sqft_basement, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft of Basement by House Quality")

dp10<-ggplot2::ggplot(train,aes(x=yr_built, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Year Built by House Quality")

dp11<-ggplot2::ggplot(train,aes(x=yr_renovated, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of View by House Quality")

dp12<-ggplot2::ggplot(train,aes(x=sqft_living15, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Living15 by House Quality")


dp13<-ggplot2::ggplot(train,aes(x=sqft_lot15, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Lots15 by House Quality")

gridExtra::grid.arrange(dp9, dp10, dp11, dp12, dp13, ncol = 2, nrow = 3)
```

```{r}
# Full model 

result_log<-glm(good_quality~. - id - date - condition - grade, family=binomial, data=train)
summary(result_log)

faraway::vif(result_log)
```

```{r}
# Reduced 

reduced_log<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15, family=binomial, data=train)
summary(reduced_log)

#  sqft_above  has high multi collinearity with sqft_living
faraway::vif(reduced_log)

#better than full model 
TS<-reduced_log$deviance-result_log$deviance
1-pchisq(TS,5)
```

```{r}
# Remove multicollinearity - compare removing sqft_above vs sqft_living

reduced_log2<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_above, family=binomial, data=train)
summary(reduced_log2)

#  sqft_above  has high multi collinearity with sqft_living
#sqft_living  44.81711 once removing sqft_above
faraway::vif(reduced_log2)
4.059e-04

# Wald test to assess one coefficient
z_stat = 0.0004059/0.00008755
2*(1-pnorm(abs(z_stat)))

```

```{r}
# Reduced Further

reduced_log3<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_living, family=binomial, data=train)
summary(reduced_log3) #higher AIC than above

#  sqft_above  has high multi collinearity with sqft_living
#sqft_above 38.06450 once removing sqft_living
faraway::vif(reduced_log3)

# Wald test to assess one coefficient
z_stat = -0.0004153/0.00008494
2*(1-pnorm(abs(z_stat)))


```

```{r}
train = train[,-c(1,2)]
test = test[,-c(1,2)]

reduced_log2<-glm(good_quality~. - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_above, family=binomial, data=train)

sapply(data, class)

preds<-predict(reduced_log2,newdata=test, type="response")

##add predicted probabilities and classification based on threshold
test.new<-data.frame(test,preds,preds>0.5)
##disply actual response, predicted prob, and classification based on threshold
head(test.new[,c(20,21,22)], )



table(test$good_quality, preds>0.5)

```














Jetts logistic regression model that predicts if a home is of good quality:
```{r}

train$waterfront = factor(train$waterfront)

chart1<-ggplot2::ggplot(train, aes(x=view, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="View", y="Proportion",
       title="Proportion of Quality by View")

chart2<-ggplot2::ggplot(train, aes(x=floors, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Floors", y="Proportion",
       title="Proportion of Quality by Floors")

chart3<-ggplot2::ggplot(train, aes(x=condition, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Condition", y="Proportion",
       title="Proportion of Quality by Condition")

chart4<-ggplot2::ggplot(train, aes(x=waterfront, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Waterfront", y="Proportion",
       title="Proportion of Quality by Waterfront")

chart1
chart2
chart3
chart4
```
```{r}
train$waterfront <- as.numeric(train$waterfront)
round(cor(train[,3:21], use= "complete.obs"),3)
```
```{r}
result<-glm(good_quality~.- date - sqft_living - id, family=binomial, data=train)
summary(result)
```

```{r}
faraway::vif(result)
```
Based on this and the summary, I will drop bathrooms, floors, view, sqft basement, yr renovated, lat, and sqft lot15 from the model to try and lower VIFs:
```{r}
result2<-glm(good_quality~.- date - sqft_living - id - view - sqft_basement - floors - bathrooms - yr_renovated - lat - sqft_lot15, family=binomial, data=train)
summary(result2)
```
```{r}
faraway::vif(result2)
```
performing a test to assess result 1 vs result 2 (reduced model):
```{r}
TS<-result2$deviance-result$deviance
TS
```
```{r}
1-pchisq(TS,7) #7 terms are dropped here
```
Since we cannot reject the null, the reduced model fits the data as well as the full model and should be used.

