```{r}
library(tidyverse)
library(MASS)
library(ggplot2)
library(faraway)

data <- read.csv("kc_house_data.csv")
data
```

# Section 3 - Data Cleanup
```{r}
non_renovated_data <- data[data$yr_renovated == 0, ]
renovated_data <- data[data$yr_renovated != 0, ]

# Katherine's Cleaning
data = data %>% 
  filter(bedrooms != 33)

data1 = data

data1$date = substring(data$date, 1,8)
data1$date = as.Date(strptime(data1$date, "%Y%m%d"))
data1$month = month(ymd(data1$date), label = TRUE, abbr = FALSE)
data1$year = year(data1$date)

Data3 = data1 %>%
  group_by(month)%>%
  summarize(avg_price = mean(price))

data = data %>%
  mutate(good_quality = case_when(
    condition > 3 & grade > 7  ~ 1 ,
    condition <= 3 | grade <= 7 ~ 0))

data$good_quality = as.factor(data$good_quality)

train_correlated_df = train[,-c(1:2, 22)]


```

```{r}
invalid_renovation_data <- data[data$yr_renovated != 0 & data$yr_renovated < data$yr_built, ]
```


# Code to split our data into test and training
```{r}
set.seed(6021)
sample.data<-sample.int(nrow(data), floor(.50*nrow(data)), replace = F)

train<-data[sample.data, ]
test<-data[-sample.data, ]
```

# Section 4 - Data Visulizations of how prices of homes are related to other variables
```{r}

ggplot(train, aes(x = sqft_living, y = price)) +
  geom_point(alpha = 0.4) +
  labs(
    title = "Scatter Plot: Price vs. Living Area",
    x = "Living Area (sqft)",
    y = "Price"
  )
```

```{r}
ggplot(train, aes(x = as.factor(condition), y = price)) +
  geom_boxplot() +
  labs(
    title = "Box Plot: Price by Condition",
    x = "Condition",
    y = "Price"
  )
```

```{r}
ggplot(train, aes(x = as.factor(condition), y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Box Plot: Price by Condition",
    x = "Condition",
    y = "Price (Log)"
  )
```

**(1) Price of median sqft living **
```{r}
train<-train%>%
  mutate(house_size=case_when(sqft_living <= 2000 ~ "small",
  sqft_living > 2000 & sqft_living <=5000 ~ "medium",
  sqft_living > 5000 ~ "large"))

medianPrice<-train %>% 
  group_by(house_size) %>% 
  summarize(median_price=median(price)) 

ggplot(medianPrice, aes(x=house_size, y=median_price))+
  geom_bar(fill="#90e7fc", stat='identity')+
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 90))+
  labs(x="Square foot living", y="Median Price of House",
  title="Price of House by Sq Ft")
```

**(2) Price of sqft living and waterfront view **
```{r}
ggplot(train, aes(x=sqft_living, y=price, color=sqft_living)) +
                 facet_wrap(~waterfront) +
                 geom_point() + 
                 labs(title = "Price of House based on sqft living and waterfront view",
                        x="Square foot living space",
                        y="Price")+ 
                theme(plot.title = element_text(hjust = 0.5))
```

**(3) Price and grade scatterplot**

```{r}
ggplot(train, aes(x=grade, y=price, color=grade)) +
                 geom_point() + 
                 labs(title = "House Price based on Grade",
                        x="Grade",
                        y="Price") + 
                theme(plot.title = element_text(hjust = 0.5))
```

**(4) Price and grade boxplot**

```{r}
ggplot(train, aes(x= as.factor(grade), y=price))+ 
  geom_boxplot()+
  theme(plot.title = element_text(hjust=0.5))+
  labs(x="Grade", y="Price", title="Price of House across Grade")
```




```{r}
####### Scatterplot of Price over Bedrooms ########### 
ggplot(train, aes(y=price,x=bedrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bedroom Count",x="Bedroom Count", y="Price ($ USD)") + geom_smooth(method=lm, se=FALSE)


####### Scatterplot of Price over Bathrooms ########### 
ggplot(train, aes(y=price,x=bathrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bathroom Count", x="Bathroom Count", y="Price ($ USD)")+ geom_smooth(method=lm, se=FALSE)

ggplot(train,aes(x=bedrooms,y=price)) + 
        geom_bar(stat = "identity", na.rm=TRUE)


########### Scatterplot of Price Over Time ########### 
ggplot(data, aes(y=price,x=date))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price Over Time",x="Time", y="Price ($ USD)")

########### Histogram Total Price by Month ########### 
#ggplot(train,aes(x=month)) + 
#        geom_bar(na.rm=TRUE)+
#         labs(title = "Histogram: Total Houses Sold Over Time",x="Time", y="Count")


########### Average Price by Month ########### 
ggplot(Data3, aes(y=avg_price,x=month))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Average Price Over Time",x="Month", y="Price ($ USD)")



```

```{r}
# Correlation for linear regression 
round(cor(train_correlated_df),3)

```

# Section 5 - Linear Regression Model Selection
```{r}
#Time was removed due to the difficulties in working with it in a linear regression model, and sqft living was also removed as it is just the sum of sqft above and sqft basement and is therefore redundant
train
result1 <- lm(price ~ . - date - id -sqft_above -sqft_basement - good_quality - zipcode -lat -long - sqft_living15, data = train)
summary(result1)

library(faraway)
round(faraway::vif(result1),3)

regfull1 <- result1


```
```{r}
#Time was removed due to the difficulties in working with it in a linear regression model, and sqft living was also removed as it is just the sum of sqft above and sqft basement and is therefore redundant

train = train[,-23]#FIX remove house_size earlier in data
  
result2 <- lm(price ~ . - date - id - sqft_above -sqft_basement - good_quality - zipcode -lat -long - sqft_living15 - sqft_lot15, data = train)
summary(result2)

result3 <- lm(price ~ . - date - id - sqft_above -sqft_basement - good_quality - zipcode -lat -long - sqft_lot, data = train)
summary(result3)


library(faraway)
round(faraway::vif(result2),3)

regfull1 <- result1


```

```{r}
result <- lm(price ~ . - date - id -sqft_above -sqft_basement - good_quality - zipcode -lat -long, data = train)
summary(result)
```

```{r}
train_automated_df = train[,-c(1:2,7,13:14,17:19,22)]

regnull <- lm(price~1, data=train_automated_df)
regfull <- result3
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```

```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction="forward")
```

```{r}
step(regnull, scope=list(lower=regnull, upper=regfull), direction="both")
```

**(Regression assumptions check)**
```{r}


par(mfrow = c(2, 2))
plot(result3)
```
```{r}
library(MASS)
MASS::boxcox(result)
```

```{r}
#logdata <- train[,-c(1:2,22:23)] # use train_automated_df instead
train_automated_df$y.star<-log(train_automated_df$price)
result.ystar<-lm(y.star~., data=train_automated_df)
par(mfrow = c(2, 2))
plot(result.ystar)
```
**(Check for multicollinearity)**
```{r}
round(faraway::vif(result),3)
```
No variables have a VIF of above 5, meaning this test does not indicate multicolinearity. sqftabove however has a 4.923 which is close enough to look into!

Note: Trying out potential other model:
```{r}
reduced <- lm(price ~ . - id - sqft_basement - date - good_quality, data = train)
summary(reduced)
```
```{r}
par(mfrow = c(2, 2))
plot(reduced)
```
```{r}
library(faraway)
round(faraway::vif(reduced),3)
```

ANOVA test comparison
```{r}
anova(reduced, result)
```
This supports the full model over the reduced model



# Section 6 - Data Visualizations to explore characteristics of good quality homes
Katherine's visualizations
```{r}
#Correlations of quantitative predictors
round(cor(train[,3:21], use= "complete.obs"),3)

train$good_quality = as.factor(train$good_quality)

dp1<-ggplot2::ggplot(train,aes(x=bedrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bedrooms by House Quality")

dp2<-ggplot2::ggplot(train,aes(x=bathrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bathrooms by House Quality")

dp3<-ggplot2::ggplot(train,aes(x=sqft_living, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Bedrooms by House Quality")

dp4<-ggplot2::ggplot(train,aes(x=floors, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Number of Floors by House Quality")

gridExtra::grid.arrange(dp1, dp2, dp3, dp4, ncol = 2, nrow = 2)
```

```{r}

dp5<-ggplot2::ggplot(train,aes(x=sqft_lot, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft of Lot by House Quality")

dp6<-ggplot2::ggplot(train,aes(x=waterfront, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Waterfron by House Quality")

dp7<-ggplot2::ggplot(train,aes(x=view, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of View by House Quality")

dp8<-ggplot2::ggplot(train,aes(x=sqft_above, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Above by House Quality")

gridExtra::grid.arrange(dp5, dp6, dp7, dp8, ncol = 2, nrow = 2)

```

```{r}
dp9<-ggplot2::ggplot(train,aes(x=sqft_basement, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft of Basement by House Quality")

dp10<-ggplot2::ggplot(train,aes(x=yr_built, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Year Built by House Quality")

dp11<-ggplot2::ggplot(train,aes(x=yr_renovated, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of View by House Quality")

dp12<-ggplot2::ggplot(train,aes(x=sqft_living15, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Living15 by House Quality")


dp13<-ggplot2::ggplot(train,aes(x=sqft_lot15, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Sq Ft. Lots15 by House Quality")

gridExtra::grid.arrange(dp9, dp10, dp11, dp12, dp13, ncol = 2, nrow = 3)
```

Jett's visualizations:
```{r}

train$waterfront = factor(train$waterfront)

chart1<-ggplot2::ggplot(train, aes(x=view, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="View", y="Proportion",
       title="Proportion of Quality by View")

chart2<-ggplot2::ggplot(train, aes(x=floors, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Floors", y="Proportion",
       title="Proportion of Quality by Floors")

chart3<-ggplot2::ggplot(train, aes(x=condition, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Condition", y="Proportion",
       title="Proportion of Quality by Condition")

chart4<-ggplot2::ggplot(train, aes(x=waterfront, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Waterfront", y="Proportion",
       title="Proportion of Quality by Waterfront")

chart1
chart2
chart3
chart4
```

Zoha's visualizations:
Quantitative Visualizations with Logistic Regression
```{r}
dp1<-ggplot2::ggplot(train,aes(x=floors, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Floors")

dp2<-ggplot2::ggplot(train,aes(x=bathrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Number of Bathrooms")

dp3<-ggplot2::ggplot(train,aes(x=bedrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Number of Bedrooms")

dp4<-ggplot2::ggplot(train,aes(x=price, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Price")

#interesting that good quality houses peaked in 1975, or does this just mean we have more data for this timeframe?
dp5<-ggplot2::ggplot(train,aes(x=yr_built, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Year Built")

dp1
dp2
dp3
dp4
dp5

gridExtra::grid.arrange(dp1, dp2, dp3, dp4, dp5, ncol = 3, nrow = 2)
```

Categorical Visualizations with Logistic Regression
```{r}
train$waterfront = factor(train$waterfront)
train$view = factor(train$view)
train$floors = factor(train$floors)
train$bathrooms = factor(train$bathrooms)
train$bedrooms = factor(train$bedrooms)

chart1<-ggplot2::ggplot(train, aes(x=waterfront, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Waterfront", y="Proportion",
  title="Proportion of Good Quality Houses by Waterfront View")

chart2<-ggplot2::ggplot(train, aes(x=view, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="View", y="Proportion",
  title="Proportion of Good Quality Houses by View")

chart3<-ggplot2::ggplot(train, aes(x=floors, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Floors", y="Proportion",
  title="Proportion of Good Quality Houses by Floors")

chart4<-ggplot2::ggplot(train, aes(x=sqft_category, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Square Foot Category", y="Proportion",
  title="Proportion of Good Quality Houses by Square Foot")

chart5<-ggplot2::ggplot(train, aes(x=bathrooms, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Bathrooms", y="Proportion",
  title="Proportion of Good Quality Houses by Number of Bathrooms")

chart5<-ggplot2::ggplot(train, aes(x=bedrooms, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Bedrooms", y="Proportion",
  title="Proportion of Good Quality Houses by Number of Bedrooms")

chart1
chart2
chart3
#chart4
chart5

```

# Section 7 - Logistic Regression Model Selection
```{r}
# Full model 

result_log<-glm(good_quality~. - id - date - condition - grade, family=binomial, data=train)
summary(result_log)

#sqft basement needs to not be NA
faraway::vif(result_log)
```

```{r}
# Reduced 

reduced_log<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15, family=binomial, data=train)
summary(reduced_log)

#  sqft_above  has high multi collinearity with sqft_living
faraway::vif(reduced_log)

#better than full model 
TS<-reduced_log$deviance-result_log$deviance
TS
#p-value
1-pchisq(TS,5)
#Critical value
qchisq(1-0.05,5)
```

```{r}
# Remove multicollinearity - compare removing sqft_above vs sqft_living

reduced_log2<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_above, family=binomial, data=train)
summary(reduced_log2)

#  sqft_above  has high multi collinearity with sqft_living
#sqft_living  44.81711 once removing sqft_above
faraway::vif(reduced_log2)
4.059e-04

# Wald test to assess one coefficient
z_stat = 0.0004059/0.00008755
2*(1-pnorm(abs(z_stat)))

```

```{r}
# Reduced Further

reduced_log3<-glm(good_quality~. - id - date - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_living, family=binomial, data=train)
summary(reduced_log3) #higher AIC than above

#  sqft_above  has high multi collinearity with sqft_living
#sqft_above 38.06450 once removing sqft_living
faraway::vif(reduced_log3)

# Wald test to assess one coefficient
z_stat = -0.0004153/0.00008494
2*(1-pnorm(abs(z_stat)))


```

```{r}
train = train[,-c(1,2)]
test = test[,-c(1,2)]

reduced_log2<-glm(good_quality~. - condition - grade - view - waterfront -  sqft_basement - sqft_lot - sqft_lot15 - sqft_above, family=binomial, data=train)

sapply(data, class)

#preds<-predict(reduced_log2,newdata=test, type="response")

##add predicted probabilities and classification based on threshold
#test.new<-data.frame(test,preds,preds>0.5)
##disply actual response, predicted prob, and classification based on threshold
#head(test.new[,c(20,21,22)], )



#table(test$good_quality, preds>0.5)

```





