```{r}
library(tidyverse)
library(MASS)
library(ggplot2)
library(faraway)

data <- read.csv("kc_house_data.csv")
data
```

# Section 3 - Data Cleanup
```{r}
non_renovated_data <- data[data$yr_renovated == 0, ]
renovated_data <- data[data$yr_renovated != 0, ]
invalid_renovation_data <- data[data$yr_renovated != 0 & data$yr_renovated < data$yr_built, ]

data = data %>% 
  filter(bedrooms != 33)

time_data = data

data$waterfront <- factor(data$waterfront)
data$condition <- factor(data$condition)
data$view <- factor(data$view)
data$grade <- factor(data$grade)

data <- data[, -which(names(data) %in% c("id", "zipcode", "lat",  "long", "date","sqft_above", "sqft_basement", "sqft_lot15","sqft_living15"))]

#Splitting our code into test and training subsets
set.seed(6021)
sample.data<-sample.int(nrow(data), floor(.50*nrow(data)), replace = F)

train<-data[sample.data, ]
test<-data[-sample.data, ]

time_data$date = substring(time_data$date, 1,8)
time_data$date = as.Date(strptime(time_data$date, "%Y%m%d"))
time_data$month = month(ymd(time_data$date), label = TRUE, abbr = FALSE)
time_data$year = year(time_data$date)

time_data_by_month = time_data %>%
  group_by(month)%>%
  summarize(avg_price = mean(price))
```

# Section 4 - Data Visulizations of how prices of homes are related to other variables
```{r}
ggplot(train, aes(x = sqft_living, y = price)) +
  geom_point(alpha = 0.4) +
  labs(
    title = "Scatter Plot: Price vs. Living Area",
    x = "Living Area (sqft)",
    y = "Price"
  )
```
```{r}
ggplot(train, aes(x = as.factor(condition), y = price)) +
  geom_boxplot() +
  scale_y_log10() +
  labs(
    title = "Box Plot: Price by Condition",
    x = "Condition",
    y = "Price (Log)"
  )
```

**(1) Price of median sqft living **
```{r}
train_house_size<-train%>%
  mutate(house_size=case_when(sqft_living <= 2000 ~ "small",
  sqft_living > 2000 & sqft_living <=5000 ~ "medium",
  sqft_living > 5000 ~ "large"))

medianPrice<-train_house_size %>% 
  group_by(house_size) %>% 
  summarize(median_price=median(price)) 

ggplot(medianPrice, aes(x=house_size, y=median_price))+
  geom_bar(fill="#90e7fc", stat='identity')+
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 90))+
  labs(x="Square foot living", y="Median Price of House",
  title="Price of House by Sq Ft")
```

**(2) Price of sqft living and waterfront view **
```{r}
ggplot(train, aes(x=sqft_living, y=price, color=sqft_living)) +
                 facet_wrap(~waterfront) +
                 geom_point() + 
                 labs(title = "Price of House based on sqft living and waterfront view",
                        x="Square foot living space",
                        y="Price")+ 
                theme(plot.title = element_text(hjust = 0.5))
```

**(3) Price and grade scatterplot**

```{r}
ggplot(train, aes(x=grade, y=price, color=grade)) +
                 geom_point() + 
                 labs(title = "House Price based on Grade",
                        x="Grade",
                        y="Price") + 
                theme(plot.title = element_text(hjust = 0.5))
```

**(4) Price and grade boxplot**

```{r}
ggplot(train, aes(x= as.factor(grade), y=price))+ 
  geom_boxplot()+
  theme(plot.title = element_text(hjust=0.5))+
  labs(x="Grade", y="Price", title="Price of House across Grade")
```
```{r}
####### Scatterplot of Price over Bedrooms ########### 
ggplot(train, aes(y=price,x=bedrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bedroom Count",x="Bedroom Count", y="Price ($ USD)") + geom_smooth(method=lm, se=FALSE)


####### Scatterplot of Price over Bathrooms ########### 
ggplot(train, aes(y=price,x=bathrooms))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Price by Bathroom Count", x="Bathroom Count", y="Price ($ USD)")+ geom_smooth(method=lm, se=FALSE)

ggplot(train,aes(x=bedrooms,y=price)) + 
        geom_bar(stat = "identity", na.rm=TRUE)

########### Average Price by Month ########### 
ggplot(time_data_by_month, aes(y=avg_price,x=month))+
         geom_point(alpha=1, shape =21)+
         labs(title = "Scatter Plot: Average Price Over Time",x="Month", y="Price ($ USD)")
```

# Section 5 - Linear Regression Model Selection
```{r}
# Correlation for linear regression 
train_without_categoricals =  train[, -which(names(train) %in% c("waterfront", "grade","view", "condition"))]
round(cor(train_without_categoricals),3)
```
```{r}
#data <- data[, -which(names(data) %in% c("waterfront", "grade","view"))]
result <- lm(price ~ ., data = train)
summary(result)

#try reducing model by removing categorical variables

reduced <- lm(price ~ ., data = train_without_categoricals)
summary(reduced)

#test VIF values
round(faraway::vif(result),3)

#Reject the null, so the data supports the full model
anova(reduced, result)

par(mfrow = c(2, 2))
plot(result)
```
```{r}
regnull <- lm(price~1, data=train)
regfull <- result
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```
```{r}
library(MASS)
MASS::boxcox(result)
```
**(Regression assumptions check)**
```{r}
logtrain <- train
logtrain$price <- log(logtrain$price)

resultlog <- lm(price ~ ., data = logtrain)
summary(resultlog)


par(mfrow = c(2, 2))
plot(resultlog)

#Test without influential point 
train_without_influential = data[-7253,]

result2 <- lm(price ~ ., data = train_without_influential)
summary(result2)


library(MASS)
MASS::boxcox(result)
```
**(Check for multicollinearity)**
```{r}
round(faraway::vif(result),3)
```

```{r}
#sample data from test data

newdata<-test[12,]

#Returns interval (-116580.5, 681017.1)
predict(resultlog, newdata, level=0.95,
        interval="prediction")

exp(12.03487)
exp(13.25534)

#Actual price 230000
test[12,1]
```

```{r}
predictions <- predict(resultlog, newdata = test)


predictions_actual_df = data.frame('predictions' = exp(predictions), 'actual' = test$price)
       
predictions_actual_df$diff_sq = (predictions_actual_df$actual - predictions_actual_df$predictions)**2

ss_res <- sum(predictions_actual_df$diff_sq)

ss_tot <- sum((predictions_actual_df$actual - mean(predictions_actual_df$actual)) **2)

r_squared <- 1-(ss_res/ss_tot)

r_squared

```


# Section 6 - Data Visualizations to explore characteristics of good quality homes
```{r}
train$condition <- as.numeric(train$condition)
train$grade <- as.numeric(train$grade)
test$condition <- as.numeric(test$condition)
test$grade <- as.numeric(test$grade)

train = train %>%
  mutate(good_quality = case_when(
    condition > 3 & grade > 7  ~ 1 ,
    condition <= 3 | grade <= 7 ~ 0))

test = test %>%
  mutate(good_quality = case_when(
    condition > 3 & grade > 7  ~ 1 ,
    condition <= 3 | grade <= 7 ~ 0))

train <- train[, -which(names(train) %in% c("condition", "grade"))]

train$good_quality = as.factor(train$good_quality)
```


```{r}
dp1<-ggplot2::ggplot(train,aes(x=floors, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Floors")

dp2<-ggplot2::ggplot(train,aes(x=bathrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Number of Bathrooms")

dp3<-ggplot2::ggplot(train,aes(x=bedrooms, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Number of Bedrooms")

dp4<-ggplot2::ggplot(train,aes(x=price, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Price")

dp5<-ggplot2::ggplot(train,aes(x=yr_built, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Year Built")

dp6<-ggplot2::ggplot(train,aes(x=yr_renovated, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Year Renovated")

dp7<-ggplot2::ggplot(train,aes(x=sqft_lot, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Sq Ft of Lot")

dp8<-ggplot2::ggplot(train,aes(x=sqft_living, color=good_quality))+
  geom_density()+
  labs(title="Density Plot of Good Quality Houses by Square foot Living")


dp1
dp2
dp3
dp4
dp5
dp6
dp7
dp8
```
```{r}
train$waterfront = factor(train$waterfront)
train$view = factor(train$view)

chart1<-ggplot2::ggplot(train, aes(x=waterfront, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Waterfront", y="Proportion",
  title="Proportion of Good Quality Houses by Waterfront View")

chart2<-ggplot2::ggplot(train, aes(x=view, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="View", y="Proportion",
  title="Proportion of Good Quality Houses by View")

chart3<-ggplot2::ggplot(train, aes(x=floors, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Floors", y="Proportion",
  title="Proportion of Good Quality Houses by Floors")

chart4<-ggplot2::ggplot(train, aes(x=sqft_category, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Square Foot Category", y="Proportion",
  title="Proportion of Good Quality Houses by Square Foot")

chart5<-ggplot2::ggplot(train, aes(x=bathrooms, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Bathrooms", y="Proportion",
  title="Proportion of Good Quality Houses by Number of Bathrooms")

chart5<-ggplot2::ggplot(train, aes(x=bedrooms, fill=good_quality))+
  geom_bar(position = "fill")+
  labs(x="Bedrooms", y="Proportion",
  title="Proportion of Good Quality Houses by Number of Bedrooms")

chart1
chart2
chart3
chart5

```


# Section 7 - Logistic Regression Model Selection
```{r}
result_log<-glm(good_quality~., family=binomial, data=train)
summary(result_log)

faraway::vif(result_log)
```

```{r}
regnull <- glm(good_quality~1, family=binomial, data=train)
regfull <- result_log
step(regfull, scope=list(lower=regnull, upper=regfull), direction="backward")
```

```{r}
# Reduced again

reduced2_log<- glm(formula = good_quality ~ price + bedrooms + sqft_living + 
    view + yr_built + yr_renovated, family = binomial, data = train)
summary(reduced2_log)


faraway::vif(reduced2_log)
```
```{r}
#better than full model 
TS<-reduced2_log$deviance-result_log$deviance
TS
#p-value
1-pchisq(TS,4)
#Critical value
qchisq(1-0.05,4)
```

The null hypothesis supports droppping the 4 predictors, and the alternate hypothesis supports not dropping the 4 predictors. We fail to reject the null hypothesis, so we drop ---- to use the reduced model. 


```{r}

preds<-predict(reduced2_log,newdata=test, type="response")

##add predicted probabilities and classification based on threshold
test.new<-data.frame(test,preds,preds>0.5)
##disply actual response, predicted prob, and classification based on threshold
head(test.new[,c(13,14,15)], )



table(test$good_quality, preds>0.5)
```

```{r}
table(test$good_quality, preds>0.56)
```

#Calculate ROC Curve

```{r}
library(ROCR)
##produce the numbers associated with classification table
rates<-ROCR::prediction(preds, test$good_quality)

##store the true positive and false positive rates
roc_result<-ROCR::performance(rates,measure="tpr", x.measure="fpr")

##plot ROC curve and overlay the diagonal line for random guessing
plot(roc_result, main="ROC Curve for Reduced Model")
lines(x = c(0,1), y = c(0,1), col="red")
#points(x=0.2068966, y=0.6393443, col="blue", pch=16)

```


```{r}
##compute the AUC
auc<-performance(rates, measure = "auc")
auc@y.values
```

